---
title: "Homework-21"
author: "S.O'Brien"
format: html
---

# set things up
```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

# import data 
```{r}
lady <- read.csv(here("Data/ladybirds.csv"))
```

# checkout data
```{r}
glimpse(lady)
```

# convert Habitat, Site, and colour to fctr
```{r}
lady$Habitat <- 
  as.factor(lady$Habitat)
levels(lady$Habitat)

lady$Site <- 
  as.factor(lady$Site)
levels(lady$Site)

lady$colour <- 
  as.factor(lady$colour)
levels(lady$colour)
```

# summarise total number for each colour and habitat
```{r}
totals <- lady |>
  group_by(Habitat, colour) |>
  summarise(total = sum(number))
totals
```

# make bar chart of total number for each colour and habitat
```{r}
ggplot(totals, aes(x = Habitat, y = total, fill = colour)) +
  geom_bar(stat = 'identity', 'position' = 'dodge')
# fill used for bar chart to specify the colour variable
# stat = 'identity' tells ggplot not to calculate anything
# position = 'dodge' tells ggplot to put two bars for each habitat
```
counts are ratio data, not summarizing distribution

# fix colors of plot to correspond to colour
```{r}
ggplot(totals, aes(x = Habitat, y = total, fill = colour)) +
  geom_bar(stat = 'identity', 'position' = 'dodge') +
  scale_fill_manual(values = c(black = "Black", red = "Red"))
# scale_fill_manual specifies colour values with a color
```

# x2 test
null hypothesis: colours are equally distributed among habitat
alt hypothesis: colours aren't equally distributed among habitat

#must convert totals to a matrix to run x2 test
```{r}
lady_mat <- xtabs(number ~ Habitat + colour, 
                  data = lady)
lady_mat
```

# run x2 test
```{r}
chisq.test(lady_mat)
```
reject null hypothesis, there is some association between colour and habitat. black colour more frequent in industrial habitats and red colour more frequent in rural habitat (x2 = 19.1, df = 1, p > 0.001)

# save x2 test
```{r}
lady.chi <- chisq.test(lady_mat)
names(lady.chi)
```
this makes row sumns, column sums, observed and expected values accessible

# import ozone data
```{r}
ozone <- read.csv(here("Data/ozone.csv"))
```

# checkout data
```{r}
glimpse(ozone)
```

# convert Garden.location and Garden.ID to fctr
```{r}
ozone$Garden.ID <- 
  as.factor(ozone$Garden.ID)
levels(ozone$Garden.ID)

ozone$Garden.location <- 
  as.factor(ozone$Garden.location)
levels(ozone$Garden.location)
```

# graph histogram for east and west
```{r}
ggplot(ozone, aes(x = Ozone)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~Garden.location, ncol = 1) + # creates two plots for east and west, ncol is # columns
  theme_bw()
```
looks like ozone levels differ for location

# calculate mean and SD of ozone levels for each location
```{r}
ozone |>
  group_by(Garden.location) |>
  summarise(mean_o = mean(Ozone),
            SD_o = sd(Ozone))
```
means look different

# t-test
```{r}
t.test(Ozone ~ Garden.location, data = ozone)
```
probability that the ozone levels aren't different at each location is very low because the 95% confidence interval doesn't include 0

# formal test
```{r}
var.test(Ozone ~ Garden.location, data = ozone)
```
don't really need to use this ever

