---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("Data/CrabAgePrediction.csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex <- 
  as.factor(crabs$Sex)
levels(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?
null hypothesis: there is no relationship between height and weight in female crabs
alt hypothesis: there is a relationship between height and weight in female crabs

1.  Extract the data for just female crabs.
```{r}
crabs_f <- crabs |>
  filter(Sex == "F")
```


2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:
predictor: height
response: weight
linear regression because ratio continuous variable and there is only one predictor variable

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

Plot data:
```{r}
ggplot(crabs_f, aes(x = Height, y = Weight)) +
  geom_point() +
  theme_bw() +
  ylab("Weight") +
  xlab("Height")
```
extreme value close to 3 for height
positive relationship from plot
slope: 70/1 = 70

simple linear regression model:
```{r}
crabs_model <- lm(Weight ~ Height, 
                data = crabs_f)
```


ggfortify()
```{r}
autoplot(crabs_model, smooth.colour = NA)
```
# residual v fitted
values not randomly plotted around 0 line and, suggest non-linear, and don't scatter along line randomly.
# normal Q-Q
values fall along line 
# scale-location
should be no pattern but most of the values fall in lower left area
# residuals v leverage
point with high influence at -25 and 0.45

anova() table
```{r}
anova(crabs_model)
```
F = 613.57 so error variance is small compared to explanatory variable 

summary() table
```{r}
summary(crabs_model)
```

conclusion:
we can reject the null hypothesis, there is very strong evidence that height is associated with wight. (F = 613.57, p = 2.2x10^-16, df = 1).

ggplot with regression line:
```{r}
ggplot(crabs_f, aes(x = Height, y = Weight)) +
  geom_point() +
  geom_smooth(method = 'lm')
  theme_bw() +
  ylab("Weight") +
  xlab("Height")
```
regression line looks influenced by extreme value and the standard error increase as height and weight increase.

## 
