---
title: "HW16"
author: "S.O'Brien"
format: html
---

## set things up
```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

## import data 
```{r}
cover <- read.csv(here("Data/donnerville_percent_cover.csv"))
```

## examaninig the data
```{r}
glimpse(cover)
summary(cover)
```

## changing Site, Plot type and cover_type into factor variables
```{r}
cover$Site <- 
  as.factor(cover$Site)
levels(cover$Site)

cover$Plot.type <- 
  as.factor(cover$Plot.type)
levels(cover$Plot.type)

cover$cover_type <- 
  as.factor(cover$cover_type)
levels(cover$cover_type)
```

## make dates into date variable 
```{r}
cover$Date <- dmy(cover$Date)
```

## error checking: fix site factor levels
```{r}
fix <- which(cover$Site == "DONNERVILLE - 1")
cover$Site[fix] <- "Donnerville 1"

fix1 <- which(cover$Site == "donnerville 1")
cover$Site[fix1] <- "Donnerville 1"

fix2 <- which(cover$Site == "donnerville 2")
cover$Site[fix2] <- "Donnerville 2"

fix3 <- which(cover$Site == "donnerville-3")
cover$Site[fix3] <- "Donnerville 3"

cover$Site <- factor(cover$Site)
```

## error checking: fix cover_type factor levels
```{r}
fix4 <- which(cover$cover_type == "bare")
cover$cover_type[fix4] <- "Bare"

fix5 <- which(cover$cover_type == "bryophtes")
cover$cover_type[fix5] <- "Bryophytes"

fix6 <- which(cover$cover_type == "vasc ")
cover$cover_type[fix6] <- "Vascular"

cover$cover_type <- factor(cover$cover_type)
```

## Determine the average percent cover by cover type (bare, bryophytes, vascular plants) for Donnerville 1
```{r}
cover %>% 
   group_by(cover_type) %>%
   summarise(mean_cover = mean(Percent))
```

## Determine the average percent vascular plant cover across all sites.
```{r}
cover |> 
  filter(cover_type == "Vascular") |>
  group_by(Site) |>
  summarise(mean_vasc_perc = mean(Percent))
```

