---
title: "HW18"
author: "S.O'Brien"
format: html
---

Set up workspace:
```{r}
rm(list=ls())  
library(tidyverse)
library(here)
```

Read in data:
```{r}
compensation <- read.csv(here("Data/compensation.csv"))
```

check out data:
```{r}
glimpse(compensation)
```


Make simple bivariate scatterplot:
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point()
```
aes: mapping b/n variables in dataset and graph feautures
data and aes always form first layer and geometrics plot the data

Interpretation:
positive, as root increase, fruit increases with it. two groups of grazed and ungrazed are also present. 

Change background:themes
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point() +
  theme_bw()
```

Increase size of points: size
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point(size = 5) + # place size in the geom_point layer
  theme_bw()
```

X-axis and y-axis labels: xlab() & ylab()
```{r}
ggplot(compensation, aes(x = Root, y = Fruit)) +
  geom_point(size = 5) + 
  theme_bw() +
  xlab("Root Biomass") +
  ylab("Fruit Production")
```

Change color of points to match levels: colour = 
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, colour = Grazing)) + # colour goes in first aes()
  geom_point(size = 5) + 
  theme_bw() +
  xlab("Root Biomass") +
  ylab("Fruit Production")
```
the two levels are grazed and ungrazed

Change shape of points to match levels: shape = 
```{r}
ggplot(compensation, aes(x = Root, y = Fruit, shape = Grazing)) + # shape goes in first aes()
  geom_point(size = 5) + 
  theme_bw() +
  xlab("Root Biomass") +
  ylab("Fruit Production")
```

Boxplots:
```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() + 
  theme_bw() +
  xlab("Grazing Treatment") +
  ylab("Fruit Production")
```

Adding more layers to boxplot:
```{r}
ggplot(compensation, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() + 
  geom_point(size = 4, colour = 'lightgrey', alpha = 0.5) +
  theme_bw() +
  xlab("Grazing Treatment") +
  ylab("Fruit Production")
```
geom_point() adds points to plot and change color, size, and transparency of points with the layers within it
alpha changes transparency 

Histograms:
can only have x-axis

Histogram of fruit production:
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram()
```
must change the binwidth (how wide each bar is) or number of bins

Changing number of bins:
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(bins = 10)
```

Changing binwidth:
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15)
```

Facets: creates seprate graphs for each level

Faceted Histograms:
```{r}
ggplot(compensation, aes(x = Fruit)) +
  geom_histogram(binwidth = 15) +
  facet_wrap(~Grazing)
```

Saving plots:
1. Export button under Plots tab saves the plot as image or pdf that can exported
2. ggsave: saves plot to a file; ggsave("plotname.png")
