---
title: "HW23"
author: "S.O'Brien"
format: html
---

# set things up
```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

# import data
```{r}
growth <- read.csv(here("Data/plant.growth.rate.csv"))
```

# checkout data
```{r}
glimpse(growth)
```

# plot data
```{r}
ggplot(growth, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  theme_bw() +
  ylab("Plant Growth Rate (mm/week)") +
  xlab("Soil Moisture Content")
```
positive slop, around 15
intercept somewhere between 15-20mm/week

# make a simple linear regression model
null: soil moisture has no effect on plant growth
alt: soil moisture has effect on plant growth
```{r}
reg_model <- lm(plant.growth.rate ~ soil.moisture.content, 
                data = growth)
```

# ggfortify()
```{r}
install.packages("ggfortify")
library(ggfortify)


autoplot(reg_model, smooth.colour = NA)
```
# residual v fitted
shows whether line is appropriate to fit data, cant have humps 
# normal Q-Q
evaluates normality, dots are residuals, want them on the line
# scale-location
evaluates assumption of equal variance, should be no pattern
# residuals v leverage
detects influential data points, dont want points w high influence 

# smooth.colour = NA
w/o this, includes a line of locally weighted regression that's unnecessary 

# anova() tables
```{r}
anova(reg_model)
```
F = 156.08, large so error variance is small compared to explanatory variable variable

# summary() table
```{r}
summary(reg_model)
```
we can reject the null hypothese (p = 2x10^-16 ), there is very strong evidence that soil moisture positively effects plant growth rate. for every increase unit in soil moisture, plant growth increases by 12.75 mm/week (y = 19.348, slope = 12.75, F = 156.08, = - 2x10^-16)

# updated ggplot w regression line
```{r}
ggplot(growth, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm')
  theme_bw() +
  ylab("Plant Growth Rate (mm/week)") +
  xlab("Soil Moisture Content")
```
this is the scatterplot but with the regression line. geom_smooth(method = 'lm') adds the regression line with the standard error as the grey shading over it
