---
title: "HW25"
author: "S.O'Brien"
format: html
---

# set things up
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

# 5.6 Analysis of Variance: one-way ANOVA
explantory variable is factor or categorical, not continuous

# import data
```{r}
daphnia <- read.csv("Data/Daphniagrowth.csv")
```
parasites alter growth rate?
parasites reduce growth rate compared to control?

# checkout data
```{r}
glimpse(daphnia)
```

# change parasite to fct
```{r}
daphnia$parasite <- 
  as.factor(daphnia$parasite)
levels(daphnia$parasite)
```

# plot data
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip() # flips x and y axis 
```
null hypothesis: no difference in mean growth rate between treatment
alt hypothesis: there is a difference in mean growth rate between treatment 

# anova model
```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

# check assumptions 
```{r}
autoplot(model_grow, smooth.colour = NA)
```
Q-Q plot: evaluates normality of the residuals, data stays tight along line

overall, assumptions look OK

# interpret output: anova()
```{r}
anova(model_grow)
```
parasite treatment has effect on mean growth rate of daphnia. large F-value is from the large variance between treatment and a small p-value is from the small within-group variance. we can reject the null hypothesis that there is no difference in mean growth rate between treatment because there is a statistical difference in mean growth rate of the parasite treatments (F = 32.325, p = 2.571 x 10^-10, df = 3). 

# R treatment contrasts
```{r}
?contr.treatment
```
contr.treatment is used to compare different treatment levels with the baseline level

# interpret output: summary()
```{r}
summary(model_grow)
```
- R orders the treatment levels in alphabetical order with control as the 'Intercept'
- 'Estimate' is the mean growth rate for Intercept (control) and the difference for each parasite's growth rate. 
- e.g M. bicuspidata mean GR - control mean GR = M. bicuspidata estimate,
0.81 - 1.21 = -0.4
- p-value tells us whether parasite GR is significantly different than control GR

# calculate mean growth rate for each treatment level
```{r}
sumdat <- daphnia |> 
  group_by(parasite) |>
  summarise(meanGR = mean(growth.rate))

sumdat
```
control = intercept 
0.81 - 1.21 = -0.4 (M. bicuspidata mean GR - control mean GR)

# raw growth rates and mean growth rates plot
```{r}
ggplot(daphnia, aes(y = growth.rate, x = parasite)) + 
  geom_point() + 
  geom_point(sumdat, aes(x = parasite, y = meanGR)) +
  coord_flip()
```
i couldn't figure out how to do the last graph, this is as far as i got after working on it for like 35 minutes. 

