---
title: "HW26"
author: "S.O'Brien"
format: html
---

# set things up
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

# import data
```{r}
pea <- read.csv(here("Data/s_r_pea_sections.csv"))
```

# checkout data
```{r}
glimpse(pea)
```

# change Treatment to fct
```{r}
pea$Treatment <- 
  as.factor(pea$Treatment)
levels(pea$Treatment)
```

# plot data
```{r}
ggplot(pea, aes(x = Treatment, y = pea_section_length_ou)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip()
```
control has the longest length at around 68.5
2% sucrose length is around -4.5 from control 
2% glucose length is around -9.5 from control
2% fructose length is around -10.5 from control
1% fructose-1% glucose is around -10.5 from control 
df = 4
df for error = 45

null hypothesis: no differences in mean length 
alt hypothesis: there is difference in mean length

# anova model
```{r}
# change reference level to Control w/ fct relevel
pea$Treatment <- 
  relevel(pea$Treatment, ref = "Control")
levels(pea$Treatmen)

model_pea <- lm(pea_section_length_ou ~ Treatment, 
                data = pea)
```

# check assumptions
```{r}
autoplot(model_pea, smooth.colour = NA)
```
assumptions look good

# interpret model
```{r}
anova(model_pea)

summary(model_pea)
```
my preliminary estimates of mean difference in pea length of sugar treatment from control look pretty accurate. from the anova model, there is strong evidence that 1% fructose-1% glucose, 2% fructose, 2% glucose, and 2% sucrose all have significantly different mean pea lengths than the control. Therefore, we can reject the null hypothesis (F = 49.368, p = 6.737x10^-16, df = 4).

# Tukey HSD test
```{r}
tukey_pea <- glht(model_pea, linfct = mcp(Treatment = "Tukey")) # tukey model 

summary(tukey_pea) # shows comparison between every treatment 

cld(tukey_pea) # shows compact letter display of tukey results
```
from the Tukey HSD test, the Control treatment has significantly different mean length compared to every sugar treatment. 2% sucrose also has significantly different mean length compared to every sugar treatment and the control. 1% fructose-1% glucose, 2% fructose, 2% glucose don't have significantly different mean pea lengths from each other but they do from the control and 2% sucrose. 

