---
title: "HW28"
author: "S.O'Brien"
format: html
---

### set things up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
library(gridExtra)
```

### load in data

```{r}
df <- read.csv(here("Data/compensation.csv"))
```

### checkout data

```{r}
glimpse(df)
summary(df)
```
variables are Root, Fruit, and Grazing. root and fruit are cotinuous and grazing is categorical. i will need to change grazing to a fct format

### change Grazing to fct

```{r}
df$Grazing <- 
  as.factor(df$Grazing)
levels(df$Grazing)
```

two levels of grazing are Grazed and Ungrazed 

### make scatterpot of Root and Fruit

```{r}
scatter <- ggplot(df, aes(x = Root, y = Fruit)) +
  geom_point() +
  theme_bw()

scatter
```

### make boxer and whiskers plot

```{r}
box <- ggplot(df, aes(x = Grazing, y = Fruit)) +
  geom_boxplot() +
  theme_bw()

box
```

### grid extra package

```{r}
# installed gridExtra package with intall.packages("gridExtra")
# loaded in gridExtra with library(gridExtra)
# gird.arrange places the two plots on the same page
grid.arrange(scatter, box, nrow = 1)
```

### changing titles

```{r}
# change axis labels 
scatter + xlab("Root Biomass") + ylab("Fruit Production")
# add figure title
scatter + ggtitle("MY TITLE")
# combine axis labels and title
scatter + labs(title = "Very Useful Title",
               x = "Root Biomass",
               y = "Fruit Production")
```

### set axis limits 

```{r}
scatter + xlim(0,20) + ylim(0,140)
# this creates x axis from 0-20 and y axis from 0-140
```

### add text to plot

```{r}
# adds text into the figure
scatter + annotate("text", x = c(6,8), y = c(105,25),
                   label = c("Text up here", "Text down here"))
```

### adjut scale of plot

```{r}
scatter + scale_x_continuous(limits = c(4,11), breaks = 4:11)
```

### select the color for grazed and ungrazed values in scatterplot 

```{r}
# grazed will be brown, ungrazed will be green
ggplot(df, aes(Root, Fruit, colour = Grazing)) +
  geom_point() +
  scale_colour_manual(values = c(Grazed = "brown", Ungrazed = "green"))
```

### transform axis to log10 and change ticks locations

```{r}
# breaks = specifies tick locations using breaks
# trans = transforms y-axis to log10
box +
  scale_y_continuous(breaks = seq(from = 10, to = 150, by = 20),
                     trans = "log10")

?scale()
```

### change theme elements of figure

```{r}
# fill changes background color, colour changes data point color
scatter +
  theme(
    panel.background = element_rect(fill = "blue", colour = "black"), # fill =  changes background color, colour = changes data point color
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "hotpink")) # colour = changes color of major grid lines 
```

### change color of axis title text 

```{r}
box +
  theme(
    axis.title.x = element_text(colour = "cornflowerblue",
                                size = rel(2)),
    axis.text.x = element_text(angle = 45, size = 13, vjust = 0.5))
```

### alters name of levels on figure

```{r}
box +
  scale_x_discrete(limits = c("Ungrazed", "Grazed"),
                   labels = c("Control", "Grazed"))
# changes ungrazed to control 
```

### customize legend

```{r}
ggplot(df, aes(Root, Fruit, colour = Grazing)) +
  geom_point() # this is the default legend

ggplot(df, aes(Root, Fruit, colour = Grazing)) +
  geom_point() +
  theme(legend.key = element_rect(fill = NA)) # this gets rid of the grey box surrounding the legend

ggplot(df, aes(Root, Fruit, colour = Grazing)) +
  geom_point() +
  theme(legend.position = "none") # this gets rid of the legend 

?theme

ggplot(df, aes(Root, Fruit, colour = Grazing)) +
  geom_point() +
  theme(legend.position = "left") # this moves legend to left side of graph

# loading ggthemes packes
library(ggthemes)

# add ggthemes to figure
ggplot(df, aes(Root, Fruit, colour = Grazing)) +
  geom_point() +
  theme(legend.key = element_rect(fill = NA)) +
  theme_fivethirtyeight()
```

